---
title: Impactos de la herbivoría por ungulados en las comunidades de plantas leñosas del Parque Natural Los Alcornocales
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

Research compendium (code and data) used for the following publication:

Rodríguez-Sánchez F, Luque-Oliva V & Jurado V. (2018) Impactos de la herbivoría por ungulados en las comunidades de plantas leñosas del Parque Natural Los Alcornocales. *Almoraima*, en prensa.

Compendium DOI: 

CITATION: 
Rodríguez-Sánchez F, Luque-Oliva V & Jurado V. (2018) Research compendium for "Impactos de la herbivoría por ungulados en las comunidades de plantas leñosas del Parque Natural Los Alcornocales" (Version 1.0.0). Zenodo. 


## Installation

In order to run the analyses you will need to install the package first:

```{r eval = FALSE}
devtools::install_github("Pakillo/exclosures-Almoraima")
```


## Usage

There is a [`makefile.R`](https://github.com/Pakillo/exclosures-Almoraima/blob/master/makefile.R) that runs each step of the analysis in the appropriate order. 


