---
title: Impactos de la herbivoría por ungulados en las comunidades de plantas leñosas del Parque Natural Los Alcornocales
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

Research compendium (code and data) used for the following publication:

Rodríguez-Sánchez F, Luque-Oliva V & Jurado V. (2018) Impactos de la herbivoría por ungulados en las comunidades de plantas leñosas del Parque Natural Los Alcornocales. *Almoraima*, en prensa.

Compendium DOI: 

CITATION: 
Rodríguez-Sánchez F, Luque-Oliva V & Jurado V. (2018) Research compendium for "Impactos de la herbivoría por ungulados en las comunidades de plantas leñosas del Parque Natural Los Alcornocales" (Version 1.0.0). Zenodo. 


## Installation

In order to run the analyses you will need to install the package first:

```{r eval = FALSE}
devtools::install_github("Pakillo/exclosures-Almoraima")
```


## Usage

There is a [`makefile.R`](https://github.com/Pakillo/exclosures-Almoraima/blob/master/makefile.R) that runs each step of the analysis in the appropriate order. 

```{r eval=FALSE}

library(exclosures)


#### READ AND PRE-PROCESS DATA #####

## Read site info
read_siteinfo("data-raw/sites_info_raw.csv")

## Read and prepare species info
read_sppinfo(sppdata = "data-raw/species_info_raw.csv")

## Read and process raw cover data
read_rawcover(rawcover = "data-raw/exclosures_cover_raw.csv", tr.length = 25)

## Read and prepare damage data
read_damage("data-raw/exclosure_damage_raw.csv")

## Prepare dataset
make_dataset()


#### EXPLORATORY ANALYSIS ####

rmarkdown::render("analyses/EDA.Rmd")

#### MANUSCRIPT ####

rmarkdown::render("manuscript/cercados_Almoraima/cercados_Almoraima.Rmd")

```



