---
title: "Exploratory Data Analysis"
author: "Francisco Rodriguez-Sanchez"
date: "`r Sys.Date()`"
output: pdf_document
geometry: margin=0.8in
---

```{r setup, include=FALSE}
library(knitr)
#knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir=normalizePath('../'))
```


```{r packages, cache=FALSE, message=FALSE}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(arm)
```


# Read dataset

```{r}
dataset <- read_csv("data/fulldata.csv")
dataset <- mutate(dataset, inside = ifelse(inside == 1, "in", "out"))
str(dataset)
summary(dataset)
head(dataset)
tail(dataset)
```

Derive some useful datasets:
```{r}
shrubs <- filter(dataset, shrub_tree == "shrub")
trees <- filter(dataset, shrub_tree == "tree")

## Only present species, and excluding bare soil
shrubs_present <- filter(shrubs, presence == 1, acronym != "Bare")
trees_present <- filter(trees, presence == 1, acronym != "Bare")
```






# Sites

```{r}
sites <- dataset %>%
  distinct(site, inside, .keep_all = TRUE) %>%
  dplyr::select(site:inside)
  
```


## Number of sites

There are `r nrow(distinct(sites, site))` sites, with 2 transects in each site (in and out of the exclosure).

```{r}
kable(sites)
```


## Map of sites

```{r fig.keep='last'}
rSDM::occmap(as.matrix(sites[, c("long", "lat")]), type = "satellite", exp = 2)
```


## Altitude of sites

```{r}
siteplots <- ggplot(sites) 
siteplots + geom_histogram(aes(altitude))
```



# Shrubs


## Shrubs: number of species present, their mean cover, and damage at each site

```{r}
shrubcover <- shrubs %>%
  group_by(site, inside) %>%
  filter(acronym == "Bare") %>%
  dplyr::select(cover) %>%
  transmute(shrubcover = 100*(50 - cover)/50)   # shrub cover in %

shrubs.summary <- shrubs_present %>%
  group_by(site, inside) %>%
  summarise(shrub_height = unique(shrub_height),
            nspp = sum(presence),
            mean_cover = round(mean(cover), 1), 
            sd_cover = round(sd(cover), 1),
            median_cover = round(median(cover), 1),
            mean_damage = round(mean(damage, na.rm = TRUE), 1)) 

shrubs.summary <- left_join(shrubcover, shrubs.summary, by = c("site", "inside"))

kable(shrubs.summary)
```


## Shrubs: quick comparison in-out

```{r}
kable(shrubs.summary %>%
        group_by(inside) %>%
        summarise(mean_shrub.cover = mean(shrubcover),
                  mean_shrub.height = mean(shrub_height),
                  mean_nspp = mean(nspp),
                  mean_cover = mean(mean_cover),
                  mean_damage = mean(mean_damage)))
```

N.B. Given the hierarchical structure of the data (obs within sites), the most correct way
of comparing average cover & damage in and out of exclosures is using mixed models.



## Is shrub cover higher within exclosures?

```{r}
shrubplots <- ggplot(shrubs.summary, aes(x = factor(inside)))
shrubplots + geom_dotplot(aes(y = shrubcover), binaxis = "y", stackdir = "center")
```

Only Barrancones has low cover within the exclosure.

For this variable (cover), bounded between 0 and 100 %, we should use beta regression, or logit-transform the data [Wharton et al. 2011](http://doi.org/10.1890/10-0340.1). Here we opt in for the latter:

```{r}
shrubs.summary$logitcover <- arm::logit(shrubs.summary$shrubcover/100)
cover.model <- lm(logitcover ~ inside, data = shrubs.summary)
summary(cover.model)
hist(resid(cover.model))
cover.sim <- sim(cover.model, n.sims = 1000)
coefs <- coef(cover.sim)
coefs[,2] <- coefs[, 1] + coefs[, 2]
coefs <- as.data.frame(invlogit(coefs))
names(coefs) <- c("inside", "outside")
apply(coefs, 2, quantile, c(0, 0.025, 0.25, 0.5, 0.75, 0.975, 1))
apply(coefs, 2, mean)
apply(coefs, 2, sd)
```

 
```{r}
coefs <- gather(coefs, "in_out", "cover", 1:2)

ggplot(coefs, aes(in_out, cover)) +
  geom_violin()
```







## Is shrub height higher within exclosures?

```{r}
shrubplots + geom_dotplot(aes(y = shrub_height), binaxis = "y", stackdir = "center")
```


```{r}
height.model <- lm(log(shrub_height) ~ inside, data = shrubs.summary)
summary(height.model)
hist(resid(height.model))
height.sim <- sim(height.model, n.sims = 1000)
coefs <- coef(cover.sim)
coefs[,2] <- coefs[, 1] + coefs[, 2]
coefs <- as.data.frame(exp(coefs))
names(coefs) <- c("inside", "outside")
apply(coefs, 2, quantile, c(0, 0.025, 0.25, 0.5, 0.75, 0.975, 1))
apply(coefs, 2, mean)
apply(coefs, 2, sd)
```

 
```{r}
coefs <- gather(coefs, "in_out", "height", 1:2)

ggplot(coefs, aes(in_out, height)) +
  geom_violin()
```





## Are there more species within exclosures?

```{r}
shrubplots + geom_dotplot(aes(y = nspp), binaxis = "y", stackdir = "center")
```


```{r}
nspp.model <- glm(nspp ~ inside, data = shrubs.summary, family = poisson)
summary(nspp.model)
hist(resid(nspp.model))
nspp.sim <- sim(nspp.model, n.sims = 1000)
coefs <- coef(nspp.sim)
coefs[,2] <- coefs[, 1] + coefs[, 2]
coefs <- as.data.frame(exp(coefs))
names(coefs) <- c("inside", "outside")
apply(coefs, 2, quantile, c(0, 0.025, 0.25, 0.5, 0.75, 0.975, 1))
apply(coefs, 2, mean)
apply(coefs, 2, sd)
```


```{r}
coefs <- gather(coefs, "in_out", "nspp", 1:2)

ggplot(coefs, aes(in_out, nspp)) +
  geom_violin()
```








## Shrubs: spp appearing in and out of exclosures

### species appearing within
```{r}
shrubs.in <- shrubs_present %>%
  filter(inside == "in") %>%
  distinct(species) %>%
  mutate(id = seq(1:nrow(.))) %>%
  dplyr::select(id, species)
kable(shrubs.in)
```

### species appearing outside
```{r}
shrubs.out <- shrubs_present %>%
  filter(inside == "out") %>%
  distinct(species) %>%
  mutate(id = seq(1:nrow(.))) %>%
  dplyr::select(id, species)
kable(shrubs.out)
```



### Shrubs: species occurring only in or out of exclosures

### only in
```{r}
in.only <- anti_join(shrubs.in, shrubs.out, by = "species") %>% 
  dplyr::select(-id) %>%
  arrange(species)
kable(in.only)
```

### only out
```{r}
out.only <- anti_join(shrubs.out, shrubs.in, by = "species") %>% 
  dplyr::select(-id) %>%
  arrange(species)
kable(out.only)
```


## Shrubs: summary of species frequency, cover, and damage

```{r}
kable(shrubs_present %>%
  group_by(species, inside) %>%
  summarise(freq = sum(presence),
            mean_cover = round(mean(cover), 1),
            sd_cover = round(sd(cover), 1),
            median_cover = round(median(cover), 1),
            mean_damage = round(mean(damage, na.rm = TRUE), 1)))
```




# Trees

## trees: number of species present, their mean cover, and damage at each site

```{r}
treecover <- trees %>%
  group_by(site, inside) %>%
  filter(acronym == "Bare") %>%
  dplyr::select(cover) %>%
  transmute(treecover = 100*(50 - cover)/50)   # tree cover in %

trees.summary <- trees_present %>%
  group_by(site, inside) %>%
  summarise(nspp = sum(presence), 
            mean_cover = round(mean(cover), 1), 
            sd_cover = round(sd(cover), 1),
            median_cover = round(median(cover), 1)) %>%
  left_join(treecover, by = c("site", "inside"))

kable(trees.summary)
```


## trees: quick comparison in-out

```{r}
kable(trees.summary %>%
  group_by(inside) %>%
  summarise(mean_nspp = mean(nspp),
            mean_treecover = mean(treecover),
            mean_cover = mean(mean_cover)))
```

N.B. Given the hierarchical structure of the data (obs within sites), the most correct way
of comparing average cover, damage, etc in and out of exclosures is using mixed models.


## trees: spp in and out

```{r}
trees.in <- trees_present %>%
  filter(inside == "in") %>%
  distinct(species) %>%
  mutate(id = seq(1:nrow(.))) %>%
  dplyr::select(id, species)
kable(trees.in)


trees.out <- trees_present %>%
  filter(inside == "out") %>%
  distinct(species) %>%
  mutate(id = seq(1:nrow(.))) %>%
  dplyr::select(id, species)
kable(trees.out)
```



### trees: species occurring only in or out of exclosures

```{r}
in.only <- anti_join(trees.in, trees.out, by = "species") %>% 
  dplyr::select(-id) %>%
  arrange(species)
kable(in.only)


out.only <- anti_join(trees.out, trees.in, by = "species") %>% 
  dplyr::select(-id) %>%
  arrange(species)
kable(out.only)
```







# Appendix: species cover and damage per site

## Shrubs
```{r}
shrubdam <- shrubs_present %>%
  dplyr::select(site, inside, species, damage) %>%
  mutate(inside = ifelse(inside == "in", "damage_in", "damage_out")) %>%
  spread(inside, damage) 

shrubcov <- shrubs_present %>%
  dplyr::select(site, inside, species, cover) %>%
  mutate(inside = ifelse(inside == "in", "cover_in", "cover_out")) %>%
  spread(inside, cover) %>%
  arrange(site, desc(cover_in)) %>%
  left_join(shrubdam, by = c("site", "species"))

kable(shrubcov)
```


## Trees
```{r}
kable(trees_present %>%
  dplyr::select(site, inside, species, cover) %>%
  mutate(inside = ifelse(inside == "in", "cover_in", "cover_out")) %>%
  spread(inside, cover) %>%
  arrange(site, desc(cover_in))) 
```


